---
title: "Assignment 1"
---

# Assignment

My first assignment has two parts.

## (a) Summary Discussion on Data Science Optimization and Career Development
This discussion explores data science, optimization techniques, and their applications in retail, transportation, and healthcare sectors. The speaker illustrates the broad impact of data science through examples from both academic and professional experiences, emphasizing the importance of interdisciplinary approaches in this evolving field.

It is highlighted that data science differs from statistics, as it extends beyond mere model development. The field necessitates engineering skills, domain expertise, and algorithmic thinking. The significance of problem definition, data collection, and preprocessing in model building is underscored. Additionally, the role of data visualization techniques in deriving meaningful insights and assessing models under real-world conditions is emphasized.

The discussion delves into how optimization algorithms are utilized for sales forecasting and inventory management in the retail sector. In particular, machine learning and big data analytics are leveraged to provide optimal stock management solutions for retailers. The speaker also presents findings from doctoral research on autonomous vehicle technology, covering traffic flow models and simulations of autonomous vehicle convoys. The potential benefits of autonomous vehicles in improving traffic efficiency and the impact of vehicle-to-vehicle communication on traffic dynamics are critically examined.

The interdisciplinary nature of data science is reaffirmed, with a strong argument that artificial intelligence will not replace data scientists in the foreseeable future. The discussion highlights the importance of human judgment in problem definition and modeling processes, emphasizing a systematic approach to data science workflows.

Finally, in the context of career development, the talk offers guidance for industrial engineers, stressing the importance of programming skills and analytical thinking. Recommendations are provided on how to enhance problem-solving abilities and navigate the evolving landscape of data-driven decision-making.

This discussion serves as a comprehensive guide for individuals working in or interested in data science and optimization, providing practical insights into the field’s challenges and future directions.

## (b) Exploring Statistical Summaries of ‘mtcars’ dataset

```{r}

data(mtcars)
print(colnames(mtcars))

compute_stats <- function(b) {
  list(
    mean = mean(b, na.rm = TRUE),
    median = median(b, na.rm = TRUE),
    variance = var(b, na.rm = TRUE),
    IQR = IQR(b, na.rm = TRUE),
    min = min(b, na.rm = TRUE),
    max = max(b, na.rm = TRUE)
  )
}


for (col in names(mtcars)) {
  if (is.numeric(mtcars[[col]])) {  
    cat("\nStatistics for:", col, "\n")
    print(compute_stats(mtcars[[col]]))
  }
}


statistics_sapply <- sapply(mtcars, compute_stats)

print("Applying compute_stats Using sapply()")

print(statistics_sapply)

statistics_apply <- apply(mtcars, MARGIN = 2, compute_stats)
print("Applying compute_stats Using apply()")
print(statistics_apply)

```

## (c) Displaying and Manipulating the ‘na_example’ Dataset

```{r}
library(dslabs)
data("na_example")
print(na_example)


na_count <- sum(is.na(na_example))
print(na_count) #Total NA values



na_indic <- which(is.na(na_example))
print(na_indic) #Indices of NA values



before_mean <- mean(na_example, na.rm = TRUE) #mean ignoring NA values
print(before_mean) #Mean before handling NA values

before_sd <- sd(na_example, na.rm = TRUE)  #standard deviation ignoring NA values
print(before_sd) #Standard Deviation before handling NA values




median_value <- median(na_example, na.rm = TRUE) # Compute the median of non-missing values


na_replaced_median <- na_example
na_replaced_median[is.na(na_replaced_median)] <- median_value # Replace NA values with the median


non_na_values <- na_example[!is.na(na_example)]
random_values <- sample(non_na_values, size = sum(is.na(na_example)), replace = TRUE) # Select a random non-missing value


na_replaced_random <- na_example
na_replaced_random[is.na(na_replaced_random)] <- random_values # Replace NA values with random values


# Version 1 (NA replaced with median)
mean_median <- mean(na_replaced_median)  #mean- NA replaced with median
print(mean_median)   #Mean after replacing NA with median
sd_median <- sd(na_replaced_median)  #standard deviation- NA replaced with median
print(sd_median)  # Standard Deviation after replacing NA with median

# Version 2 (NA replaced with random values)
mean_random <- mean(na_replaced_random)
print(mean_random) #Mean after replacing NA with random values
sd_random <- sd(na_replaced_random)
print(sd_random) #Standard Deviation after replacing NA with random values
```

**Comparison of Results**

|                                  | Mean  | Standard Deviation |
|----------------------------------|-------|--------------------|
| Before Handling NAs              | 2.302 | 1.223              |
| Replacing NAs with Median        | 2.258 | 1.136              |
| Replacing NAs with Random Values | 2.304 | 1.223              |
